\bf An Unitary RG-based Auxiliary Model Approach\\ to Strongly-Correlated Systems

### 2023-03-06 notes/notes.tex
added 1/Z factors

@@

	\ket{\Psi_\text{gs}} = \frac{1}{\lambda_{\vec k_0} N}\sum_{\vec R_j} e^{i \vec{k}_0\cdot\vec{R}_j} \ket{\psi_\text{aux}(\vec R_j)}~.
We again simplify the wavefunction by keeping only the uniform piece and dropping the difference pieces:
	\ket{\Psi_\text{gs}} = \frac{1}{\lambda_{\vec k_0} N}\sum_{\vec R_j} e^{i \vec{k}_0\cdot\vec{R}_j} \ket{\psi_\text{aux}(\vec R_j)} = \sum_{\vec a \in \text{NN of}\vec R_0}\frac{1}{\mathcal{Z}\lambda_{\vec k_0}} e^{i\vec{k}_0\cdot\vec{R}_0} \ket{\psi_\text{aux}(\vec R_0, \vec a)} + \ket{\psi_\text{diff}(\vec k_0)}
where \(\ket{\psi_\text{diff}(\vec k_0)} = \ket{\Psi_\text{gs}} - \frac{1}{\lambda_{\vec k_0}} e^{i\vec{k}_0\cdot\vec{R}_0} \ket{\psi_\text{aux}(\vec R_0)} = \frac{1}{\lambda_{\vec k_0} N}\sum_{\vec R_j} \left[e^{i \vec{k}_0\cdot\vec{R}_j} \ket{\psi_\text{aux}(\vec R_j)} - e^{i\vec{k}_0\cdot\vec{R}_0} \ket{\psi_\text{aux}(\vec R_0)}\right]\), and we have brought back the explicit sum over the nearest neighbours \(\vec a\) of the impurity site (following eq.~\ref{zeroth_site_average}).
By retaining only the auxiliary model state with the impurity at \(\vec R_0\) and the , the Greens function can be expressed as
	G_\text{H-H}(\vec r, t) &= -\frac{1}{\lambda_{\vec k_0}^2}i\theta(t) \braket{\psi_\text{aux}(\vec R_0) | \left\{ e^{i N \mathcal{H}_\text{aux}(\vec R_0) t}c(\vec R + r)e^{-i N \mathcal{H}_\text{aux}(\vec R_0) t}, c^\dagger(\vec R) \right\} | \psi_\text{aux}(\vec R_0)}~.
	G_\text{H-H}(\vec r, t) &= -\frac{1}{\lambda_{\vec k_0}^2}i\theta(t) \braket{\psi_\text{aux}(\vec R_0) | \left\{ e^{i N \mathcal{H}_\text{aux}(\vec R_0) t}c(\vec R_0 + r)e^{-i N \mathcal{H}_\text{aux}(\vec R_0) t}, c^\dagger(\vec R_0) \right\} | \psi_\text{aux}(\vec R_0)}~.
In this form, the only Hamiltonian involved in the expression is the COM Hamiltonian at \(\vec R_0\), and \(G_\text{H-H}(\vec r, t)\) is therefore captured by the auxiliary model Greens function \(\mathcal{G}\) for the propagation of an excitation between the impurity site \(\vec R_0\) and a bath site at a distance \(\vec r\) from the impurity:

@@

	G_\text{H-H}(\vec r, t) &= \frac{1}{\lambda_{\vec k_0}^2}\mathcal{G}(d,d+\vec r;t)~.
The cursive script symbol \(\mathcal{G}\) has been used to denote the Greens function for the auxiliary model \(\mathcal{H}\).

@@

The local Greens function \((\vec r=0)\)of the full model can thus be computed from the impurity local Greens function \(\mathcal{G}_{dd}\) of the auxiliary model. The nearest neighbour Greens function is given by the auxiliary model Greens function \(\mathcal{G}_{dz}\) between the impurity site and the bath zeroth site coupled to it. By using these real-space Greens functions, one can then compute \(k-\)space Greens functions of the full model. Moreover, this approach holds for any \(n-\)-particle Greens function; we just need to compute the corresponding correlator in the auxiliary model.

@@

The momentum space Greens function can be expressed as a Fourier transform of the real space Greens functions:
\begin{equation}\begin{aligned}
	G_{H-H} (\vec k, \omega) = \frac{1}{N}\sum_{\vec r, \vec r_j}e^{i \vec{k}\cdot\left(\vec{r_i} - \vec {r_j}\right)}G_{H-H} (|\vec{r_i} - \vec {r_j}|, \omega)
\end{aligned}\end{equation}
where $\vec{r_i}$ is the position vector of a particular lattice site. Because of translation invariance, the real space Greens function depends only on the relative vector between any two sites. As a result, $\vec r=0$ gives the local Greens function, $|\vec r|=a$ gives the nearest-neighbour Greens function and so on ($a$ being the lattice spacing). Given that our auxiliary model shows an emergent self-consistency on the impurity site and on the link between the impurity and the zeroth site, we approximate the \(k-\)space Greens functions from the local and nearest-neighbour contributions, by usingthe mapping between the bulk Greens function \(G\) and auxiliary model Greens functions \(\mathcal{G}\)(eq.~\ref{greens_func_relation}):
\begin{equation}\begin{aligned}
	G_{H-H} (\vec k, \omega) \simeq \frac{1}{N}\sum_{\vec{r_i} = \vec {r_j}}\left(G_{H-H} (\omega)\right) _\text{loc} + \frac{1}{N}\sum_{|\vec{r_i} - \vec {r_j}|=a} e^{i \vec{k}\cdot\left(\vec{r_i} - \vec {r_j}\right)}\left(G_{H-H} (\omega)\right)_\text{n-n}
\end{aligned}\end{equation}
The first summation produces a factor of $N$, while the second summation can be factorized into a sum over all sites (which again returns $N$) and a sum over all the primitive vectors connecting any single site with all its nearest neighbours, $\left\{ \vec a_i: i \in \left[1, w\right]\right\}$.
\begin{equation}\begin{aligned}
	\label{k_Gf_siam}
	G_{H-H} (\vec k, \omega) &\simeq G_{H-H} (\omega)_\text{loc} + G_{H-H} (\omega)_\text{n-n}\sum_{i=1}^w e^{i \vec{k}\cdot\vec {a_i}} = \frac{1}{\lambda_{\vec k_0}^2}\left[\mathcal{G}_{dd}(\omega) + \mathcal{G}_{dz}(\omega) \xi_{\vec k}\right]~,
\end{aligned}\end{equation}
where we defined \(\xi_{\vec k} \equiv \sum_{i=1}^w e^{i \vec{k}\cdot\vec {a_i}}\), it being essentially the disperson in units of the hopping strength \(t\). For example, on a d-dimensional hypercubic lattice, we obtain
\begin{equation}\begin{aligned}
	\xi_{\vec k} = \sum_{i=1}^d \left(e^{i k_i {a_i}} + e^{-i k_i {a_i}}\right) = 2\sum_{i=1}^d \cos k_i a_i~,
\end{aligned}\end{equation}
leading to \(2(\cos q_{x}a + \cos q_{y}a)\equiv \frac{-\epsilon_{\vec{q}}}{t^{H}}\) on a 2D square lattice.

The spectral function is simply the imaginary part of the Greens function: \(A_\text{H-H} = -\frac{1}{\pi}\text{Im}\left[G_{H-H}\right] \); as a result, they have similar relations with their auxiliary model counterparts:
\begin{equation}\begin{aligned}
	\left(A_\text{H-H}\right)_\text{loc}(\omega) = \frac{1}{\lambda_{\vec k_0}^2}\mathcal{A}_{dd}~,\\
	A_\text{H-H}(\vec{k},\omega) = \frac{1}{\lambda_{\vec k_0}^2}\left[\mathcal{A}_{dd} + \xi_{\vec k} \mathcal{A}_{dz} \right]~,
\end{aligned}\end{equation}
where \(\left(A_\text{H-H}\right)_\text{loc}(\omega)\) and \(A_\text{H-H}(\vec{k},\omega)\) are the local and \(k-\)space spectral functions, respectively, of the bulk model, and \(\mathcal{A}_{dd}\) and \(\mathcal{A}_{dz}\) are the auxiliary model spectral functions for local and nearest neighbour fluctuations respectively.

>>>>

	G_\text{H-H}(\vec r, t) = -i\theta(t) \braket{\Psi_\text{gs} | \left\{ e^{i N \mathcal{H}_\text{aux}(\vec R_0) t}c(\vec R_0 + r)e^{-i N \mathcal{H}_\text{aux}(\vec R_0) t}, c^\dagger(\vec R_0) \right\} | \Psi_\text{gs}}~.

@@

	\ket{\Psi_\text{gs}} = \frac{1}{\lambda_{\vec k_0} N}\sum_{\vec R_j, \vec a \in \text{NN of }\vec R_j} e^{i \vec{k}_0\cdot\vec{R}_j} \ket{\psi_\text{aux}(\vec R_j, \vec a)}~,
where we have brought back the explicit sum over the nearest neighbours \(\vec a\) of the impurity site (following eq.~\ref{zeroth_site_average}).
We again simplify the wavefunction by keeping only the uniform piece and dropping the difference pieces.
	\ket{\Psi_\text{gs}} = \sum_{\vec a \in \text{NN of}\vec R_0}\frac{1}{\mathcal{Z}\lambda_{\vec k_0}} e^{i\vec{k}_0\cdot\vec{R}_0} \ket{\psi_\text{aux}(\vec R_0, \vec a)} + \ket{\psi_\text{diff}(\vec k_0)}
where \(\ket{\psi_\text{diff}(\vec k_0)} = \ket{\Psi_\text{gs}} - \frac{1}{\lambda_{\vec k_0}} e^{i\vec{k}_0\cdot\vec{R}_0} \ket{\psi_\text{aux}(\vec R_0)} = \frac{1}{\lambda_{\vec k_0} N}\sum_{\vec R_j} \left[e^{i \vec{k}_0\cdot\vec{R}_j} \ket{\psi_\text{aux}(\vec R_j)} - e^{i\vec{k}_0\cdot\vec{R}_0} \ket{\psi_\text{aux}(\vec R_0)}\right]\).
Specifically, for the local Greens function, we keep the states with \(\vec R_j = \vec R_0\), leading to the expression
\begin{equation}\begin{aligned}\label{greens_func_approx}
	G_\text{H-H}(\vec r, t) &= -\frac{1}{\lambda_{\vec k_0}^2 \mathcal{Z}^2}i\theta(t) \sum_{\vec a_1, \vec a_2}\braket{\psi_\text{aux}(\vec R_0, \vec a_2) | \left\{ e^{i N \mathcal{H}_\text{aux}(\vec R_0) t}c(\vec R_0 + r)e^{-i N \mathcal{H}_\text{aux}(\vec R_0) t}, c^\dagger(\vec R_0) \right\} | \psi_\text{aux}(\vec R_0, \vec a_1)}~.
Because of the symmetry of the lattice, all Greens functions obtained for various positions of the zeroth site \(\vec a_1\) and \(\vec a_2\) are equal to their mean:
	G_\text{H-H}(\vec r=0, t) &= -\frac{1}{\lambda_{\vec k_0}^2 }i\theta(t) \braket{\psi_\text{aux}(\vec R_0) | \left\{ e^{i N \mathcal{H}_\text{aux}(\vec R_0) t}c(\vec R_0 + r)e^{-i N \mathcal{H}_\text{aux}(\vec R_0) t}, c^\dagger(\vec R_0) \right\} | \psi_\text{aux}(\vec R_0)}~.
In this form, the only Hamiltonian involved in the expression is the COM Hamiltonian at \(\vec R_0\), and \(G_\text{H-H}(\vec r=0, t)\) is therefore captured by the impurity Greens function \(\mathcal{G}\) in the auxiliary model \(N \mathcal{H}_\text{aux}(\vec R_0)\):

@@

	G_\text{H-H}(\vec r=0, t) &= \frac{1}{\lambda_{\vec k_0}^2}\mathcal{G}_{dd}(t)~.
The cursive script symbol \(\mathcal{G}_{dd}\) has been used to denote the impurity Greens function for the auxiliary model \(N \mathcal{H}_\text{aux}(\vec R_0)\).

For the nearest neighbour Greens function \(G_\text{H-H}(\vec r=\vec a, t)\), we keep only the terms with \(\vec a_1 = \vec a_2\) in the summation of eq.~\ref{greens_func_approx}:
\begin{equation}\begin{aligned}
	G_\text{H-H}(\vec r, t) &= -\frac{1}{\lambda_{\vec k_0}^2 \mathcal{Z}^2}i\theta(t) \sum_{\vec a_1}\braket{\psi_\text{aux}(\vec R_0, \vec a_1) | \left\{ e^{i N \mathcal{H}_\text{aux}(\vec R_0) t}c(\vec R_0 + r)e^{-i N \mathcal{H}_\text{aux}(\vec R_0) t}, c^\dagger(\vec R_0) \right\} | \psi_\text{aux}(\vec R_0, \vec a_1)}~.
\end{aligned}\end{equation}
Again, because of the lattice symmetry, all values of \(\vec a_1\) lead to the same impurity-bath Greens function \(\mathcal{G}_{dz}\) on the auxiliary model:
\begin{equation}\begin{aligned}
	G_\text{H-H}(\vec r, t) &= -\frac{1}{\lambda_{\vec k_0}^2 \mathcal{Z}}i\theta(t) \braket{\psi_\text{aux}(\vec R_0, \vec a) | \left\{ e^{i N \mathcal{H}_\text{aux}(\vec R_0) t}c(\vec R_0 + r)e^{-i N \mathcal{H}_\text{aux}(\vec R_0) t}, c^\dagger(\vec R_0) \right\} | \psi_\text{aux}(\vec R_0, \vec a)} = \frac{1}{\lambda_{\vec k_0}^2 \mathcal{Z}} \mathcal{G}_{dz}(t)~.
\end{aligned}\end{equation}

@@

The local and nearest-neighbour Greens functions of the full model can thus be computed from the impurity local Greens function \(\mathcal{G}_{dd}\) and the impurity-bath zeroth site Greens function \(\mathcal{G}_{dz}\) on the COM auxiliary model \(N \mathcal{H}_\text{aux}(\vec R_0)\). By using these real-space Greens functions, one can then compute \(k-\)space Greens functions of the full model. Moreover, this approach holds for any \(n-\)-particle Greens function; we just need to compute the corresponding correlator in the auxiliary model.

@@

The momentum space Greens function can be expressed as a Fourier transform of the real space Greens functions:
\begin{equation}\begin{aligned}
	G_{H-H} (\vec k, \omega) = \frac{1}{N}\sum_{\vec r, \vec r_j}e^{i \vec{k}\cdot\left(\vec{r_i} - \vec {r_j}\right)}G_{H-H} (|\vec{r_i} - \vec {r_j}|, \omega)
\end{aligned}\end{equation}
where $\vec{r_i}$ is the position vector of a particular lattice site. Because of translation invariance, the real space Greens function depends only on the relative vector between any two sites. As a result, $\vec r=0$ gives the local Greens function, $|\vec r|=a$ gives the nearest-neighbour Greens function and so on ($a$ being the lattice spacing). Given that our auxiliary model shows an emergent self-consistency on the impurity site and on the link between the impurity and the zeroth site, we approximate the \(k-\)space Greens functions from the local and nearest-neighbour contributions, by usingthe mapping between the bulk Greens function \(G\) and auxiliary model Greens functions \(\mathcal{G}\)(eq.~\ref{greens_func_relation}):
\begin{equation}\begin{aligned}
	G_{H-H} (\vec k, \omega) \simeq \frac{1}{N}\sum_{\vec{r_i} = \vec {r_j}}\left(G_{H-H} (\omega)\right) _\text{loc} + \frac{1}{N}\sum_{|\vec{r_i} - \vec {r_j}|=a} e^{i \vec{k}\cdot\left(\vec{r_i} - \vec {r_j}\right)}\left(G_{H-H} (\omega)\right)_\text{n-n}
\end{aligned}\end{equation}
The first summation produces a factor of $N$, while the second summation can be factorized into a sum over all sites (which again returns $N$) and a sum over all the primitive vectors connecting any single site with all its nearest neighbours, $\left\{ \vec a_i: i \in \left[1, \mathcal{Z}\right]\right\}$.
\begin{equation}\begin{aligned}
	\label{k_Gf_siam}
	G_{H-H} (\vec k, \omega) &\simeq G_{H-H} (\omega)_\text{loc} + G_{H-H} (\omega)_\text{n-n}\sum_{i=1}^{\mathcal{Z}} e^{i \vec{k}\cdot\vec {a_i}} = \frac{1}{\lambda_{\vec k_0}^2}\left[\mathcal{G}_{dd}(\omega) + \frac{1}{\mathcal{Z}}\mathcal{G}_{dz}(\omega) \xi_{\vec k}\right]~,
\end{aligned}\end{equation}
where we defined \(\xi_{\vec k} \equiv \sum_{i=1}^w e^{i \vec{k}\cdot\vec {a_i}}\), it being essentially the disperson in units of the hopping strength \(t\). For example, on a d-dimensional hypercubic lattice, we obtain
\begin{equation}\begin{aligned}
	\xi_{\vec k} = \sum_{i=1}^d \left(e^{i k_i {a_i}} + e^{-i k_i {a_i}}\right) = 2\sum_{i=1}^d \cos k_i a_i~,
\end{aligned}\end{equation}
leading to \(2(\cos q_{x}a + \cos q_{y}a)\equiv \frac{-\epsilon_{\vec{q}}}{t^{H}}\) on a 2D square lattice.

The spectral function is simply the imaginary part of the Greens function: \(A_\text{H-H} = -\frac{1}{\pi}\text{Im}\left[G_{H-H}\right] \); as a result, they have similar relations with their auxiliary model counterparts:
\begin{equation}\begin{aligned}
	\left(A_\text{H-H}\right)_\text{loc}(\omega) = \frac{1}{\lambda_{\vec k_0}^2}\mathcal{A}_{dd}~,\\
	A_\text{H-H}(\vec{k},\omega) = \frac{1}{\lambda_{\vec k_0}^2}\left[\mathcal{A}_{dd} + \xi_{\vec k} \frac{1}{\mathcal{Z}^2}\mathcal{A}_{dz} \right]~,
\end{aligned}\end{equation}
where \(\left(A_\text{H-H}\right)_\text{loc}(\omega)\) and \(A_\text{H-H}(\vec{k},\omega)\) are the local and \(k-\)space spectral functions, respectively, of the bulk model, and \(\mathcal{A}_{dd}\) and \(\mathcal{A}_{dz}\) are the auxiliary model spectral functions for local and nearest neighbour fluctuations respectively.
